---
title: "RClub HW1"
author: "Esteban"
date: "April 12, 2019"
output: 
  html_document: 
    keep_md: yes
---

Assignment:  
2E1, 2E2, 2E3  
2M1, 2M2, 2M3, 2M4, 2M5  
  
2E1. Which of the expressions below correspond to the statement: the probability of rain on Monday?  
(1) Pr(rain)  
(2) Pr(rain|Monday)  
**(3) Pr(Monday|rain)**  
(4) Pr(rain; Monday) = Pr(Monday)  
  
2E2. Which of the following statements corresponds to the expression: Pr(Monday|rain)?  
(1) The probability of rain on Monday.  
**(2) The probability of rain, given that it is Monday.**  
(3) The probability that it is Monday, given that it is raining.  
(4) The probability that it is Monday and that it is raining.  
  
2E3. Which of the expressions below correspond to the statement: the probability that it is Monday,
given that it is raining?  
**(1) Pr(Monday|rain)**  
(2) Pr(rain|Monday)  
(3) Pr(rain|Monday) Pr(Monday)  
(4) Pr(rain|Monday) Pr(Monday) = Pr(rain)  
(5) Pr(Monday|rain) Pr(rain) = Pr(Monday)  

---

2M1. Recall the globe tossing model from the chapter. Compute and plot the grid approximate
posterior distribution for each of the following sets of observations. In each case, assume a uniform
prior for p.  
(1) W, W, W  
(2) W, W, W, L  
(3) L, W, W, L, W, W, W  

```{r}
## (1)
# define grid
p_grid <- seq( from=0 , to=1 , length.out=20 )

# define prior
prior <- rep( 1 , 20 )

# compute likelihood at each value in grid
likelihood <- dbinom( 3 , size=3 , prob=p_grid )

# compute product of likelihood and prior
unstd.posterior <- likelihood * prior

# standardize the posterior, so it sums to 1
posterior1 <- unstd.posterior / sum(unstd.posterior)

# plot
plot( p_grid , posterior1 , type="b" , xlab="Probability of Water" , ylab="Posterior Probability" ) + mtext( "(1) W, W, W" )

## (2)
likelihood <- dbinom( 3 , size=4 , prob=p_grid )
unstd.posterior <- likelihood * prior
posterior2 <- unstd.posterior / sum(unstd.posterior)
plot( p_grid , posterior2 , type="b" , xlab="Probability of Water" , ylab="Posterior Probability" ) + mtext( "(2) W, W, W, L" )


##(3)
likelihood <- dbinom( 5 , size=7 , prob=p_grid )
unstd.posterior <- likelihood * prior
posterior3 <- unstd.posterior / sum(unstd.posterior)
plot( p_grid , posterior3 , type="b" , xlab="Probability of Water" , ylab="Posterior Probability" ) + mtext( "(3) L, W, W, L, W, W, W" )
```

?
2M2. Now assume a prior for p that is equal to zero when p < 0:5 and is a positive constant when
p > 0:5. Again compute and plot the grid approximate posterior distribution for each of the sets of
observations in the problem just above.
```{r}
## (1)
# define grid
p_grid <- seq( from=0 , to=1 , length.out=20 )

# define prior
prior2 <- rep( 1 , 20 )
prior2 <- ifelse( p_grid < 0.5 , 0 , 1 )

# compute likelihood at each value in grid
likelihood <- dbinom( 3 , size=3 , prob=p_grid )

# compute product of likelihood and prior
unstd.posterior <- likelihood * prior2

# standardize the posterior, so it sums to 1
posterior2.1 <- unstd.posterior / sum(unstd.posterior)

# plot
plot( p_grid , posterior2.1 , type="b" , xlab="Probability of Water" , ylab="Posterior Probability" ) + mtext( "(1) W, W, W" )

## (2)
likelihood <- dbinom( 3 , size=4 , prob=p_grid )
unstd.posterior <- likelihood * prior2
posterior2.2 <- unstd.posterior / sum(unstd.posterior)
plot( p_grid , posterior2.2 , type="b" , xlab="Probability of Water" , ylab="Posterior Probability" ) + mtext( "(2) W, W, W, L" )


##(3)
likelihood <- dbinom( 5 , size=7 , prob=p_grid )
unstd.posterior <- likelihood * prior2
posterior2.3 <- unstd.posterior / sum(unstd.posterior)
plot( p_grid , posterior2.3 , type="b" , xlab="Probability of Water" , ylab="Posterior Probability" ) + mtext( "(3) L, W, W, L, W, W, W" )
```


?
2M3. Suppose there are two globes, one for Earth and one for Mars. Th e Earth globe is 70% covered
in water. Th e Mars globe is 100% land. Further suppose that one of these globes-you don't know
which-was tossed in the air and produced a "land" observation. Assume that each globe was equally
likely to be tossed. Show that the posterior probability that the globe was the Earth, conditional on
seeing "land" (Pr(Earth|land)), is 0.23.

```{r}
prior.twoglobes = c(0.5, 0.5)
prob.twoglobes = c(0.7, 0)
likelihood.twoglobes = dbinom( 0, 1, prob.twoglobes)
unstd.posterior.twoglobes = likelihood.twoglobes * prior.twoglobes
posterior.twoglobes = unstd.posterior.twoglobes / sum(unstd.posterior.twoglobes)
posterior.twoglobes

```

?
2M4. Suppose you have a deck with only three cards. Each card has two sides, and each side is either
black or white. One card has two black sides. Th e second card has one black and one white side. The
third card has two white sides. Now suppose all three cards are placed in a bag and shuffl ed. Someone
reaches into the bag and pulls out a card and places it fl at on a table. A black side is shown facing up,
but you don't know the color of the side facing down. Show that the probability that the other side is
also black is 2/3. Use the counting method (Section 2 of the chapter) to approach this problem. This
means counting up the ways that each card could produce the observed data (a black side facing up
on the table).

> B/B, B/W, W/W  
> 6 sides...  
> Three ways to get black on first (2 for B/B, 1 for BW)  
> Two out of three of those ways produce B/B  

?
2M5. Now suppose there are four cards: B/B, B/W, W/W, and another B/B. Again suppose a card is
drawn from the bag and a black side appears face up. Again calculate the probability that the other
side is black.

> Five ways to get black on first...
> 4/5 of those yield B/B

