---
title: "Chapter 6 HW"
author: "Esteban"
date: "May 29, 2019"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(rethinking)
library(tidyverse)
library(dagitty)

data(WaffleDivorce)
d <- WaffleDivorce
```

## 6H1
```{r}
divorce_dag <- dagitty( "dag {
    S -> A -> D
    S -> M -> D
    S -> W -> D
    A -> M
}")
coordinates( divorce_dag ) <- list( x=c(S=0, W=1, M=0.5, A=0, D=1),
                                  y=c(S=0, W=0, M=0.5,   A=1, D=1) )
plot(divorce_dag)
```
```{r}
d$D <- scale(d$Divorce)
d$W <- scale(d$WaffleHouses)
d$S <- scale((d$South +1))
d$A <- scale(d$MedianAgeMarriage)
d$M <- scale(d$Marriage)

divorce_modS <- quap(
  alist(
    D ~ dnorm( mean=mu, sd=sigma ),
    mu <- a + bW*W + bS*S,
    a ~ dnorm( 0 , 0.2 ) ,
        bW ~ dnorm( 0 , 0.5 ) ,
        bS ~ dnorm( 0 , 0.5 ) ,
        sigma ~ dexp( 1 )
    ) , data = d )
precis(divorce_modS)
```

## 6H2
```{r}
impliedConditionalIndependencies(divorce_dag)

divorce_mod1 <- quap(
  alist(
    A ~ dnorm( mean=mu, sd=sigma ),
    mu <- a + bS*S + bW*W,
    a ~ dnorm( 0 , 0.2 ) ,
        c(bW, bS) ~ dnorm( 0 , 0.5 ) ,
        sigma ~ dexp( 1 )
    ) , data = d )
precis(divorce_mod1)

divorce_mod2 <- quap(
  alist(
    D ~ dnorm( mean=mu, sd=sigma ),
    mu <- a + bS*S + bA*A + bM*M + bW*W,
    a ~ dnorm( 0 , 0.2 ) ,
        c(bW, bS, bA, bM) ~ dnorm( 0 , 0.5 ) ,
        sigma ~ dexp( 1 )
    ) , data = d )
precis(divorce_mod2)

divorce_mod3 <- quap(
  alist(
    M ~ dnorm( mean=mu, sd=sigma ),
    mu <- a + bW*W + bS*S ,
    a ~ dnorm( 0 , 0.2 ) ,
        c(bW, bS) ~ dnorm( 0 , 0.5 ),
        sigma ~ dexp( 1 )
    ) , data = d )
precis(divorce_mod3)


```
## Question 1
1. Use a model to infer the total causal influence of area on weight. Would
increasing the area available to each fox make it heavier (healthier)? You
might want to standardize the variables. Regardless, use prior predictive
simulation to show that your model’s prior predictions stay within the pos-
sible outcome range.
```{r}
data(foxes)
f <- foxes

fox_dag <- dagitty( "dag{
                    area -> avgfood -> weight
                    avgfood -> groupsize -> weight
                    }")
coordinates( fox_dag ) <- list( x=c(area=0.5, avgfood=0, groupsize=1, weight=0.5),
                                  y=c(area=0, avgfood=0.5, groupsize=0.5, weight=1) )

plot(fox_dag)

adjustmentSets( fox_dag, exposure="area", outcome="weight" )
```
There are no back-door paths from area to weight.

```{r}
f$W <- standardize(f$weight)
f$A <- standardize(f$area)

fox_mod <- quap(
  alist(
    W ~ dnorm( mean=mu, sd=sigma ),
    mu <- a + bA*A,
    a ~ dnorm( 0 , 0.2 ) ,
        bA ~ dnorm( 0 , 0.5 ) ,
        sigma ~ dexp( 1 )
    ) , data = f )
precis(fox_mod)
plot(coeftab(fox_mod))

# plot prior parameters
par(mfrow=c(3,1))
curve( dnorm( x, 0 , 0.2 ), from=-1, to=1, ylab="prior for a" )
curve( dnorm( x, 0 , 0.5 ), from=-2, to=2, ylab="prior for bA" )
curve( dexp( x, 1), from=-0.01, to=5, ylab="prior for sigma" )

# prior predictive simulation
N <- 200
sim.a <- rnorm( N, 0, 0.2 ) 
sim.bA <- rnorm( N, 0, 0.5 )
sim.A <- seq( from=min(f$A) , to=max(f$A), length.out = N )
sim.mu <- sim.a + sim.bA * sim.A
sim.sigma <- rexp(N, 1) 
prior.W <- rnorm(N, sim.mu, sim.sigma)

par(mfrow=c(1,1))
plot( NULL , xlim=range(f$A) , ylim=c(-4,4) , xlab="area" , ylab="weight" )
for ( i in 1:N ) curve( sim.a[i] + sim.bA[i]*x ,
  from=min(f$A) , to=max(f$A) , add=TRUE ,
  col=col.alpha("black",0.2) )

dens( prior.W)

```

## Question 2
```{r}
f$F <- standardize(f$avgfood)

fox_mod2 <- quap(
  alist(
    W ~ dnorm( mean=mu, sd=sigma ),
    mu <- a + bF*F,
    a ~ dnorm( 0 , 0.2 ) ,
        bF ~ dnorm( 0 , 0.5 ) ,
        sigma ~ dexp( 1 )
    ) , data = f )
precis(fox_mod2)
plot(coeftab(fox_mod2))

```

## Question 3
```{r}
f$G <-  standardize(f$groupsize)
fox_mod3 <- quap(
  alist(
    W ~ dnorm( mean=mu, sd=sigma ),
    mu <- a + bF*F + bG*G,
    a ~ dnorm( 0 , 0.2 ) ,
        c( bF, bG ) ~ dnorm( 0 , 0.5 ) ,
        sigma ~ dexp( 1 )
    ) , data = f )
precis(fox_mod3)
plot(coeftab(fox_mod3))
```

