---
title: "Chapter 4"
author: "Esteban"
date: "April 26, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(rethinking)
```

```{r, include=FALSE}
par(mfrow=c(1,2))
pos <- replicate( 1000 , sum( runif(16,-1,1) ) )
plot(density(pos))
hist(pos) 
curve(exp( -x^2 ), from=-3, to=3)
dnorm(0,0,0.1)
```



Reading: Chapter 4 through section 4.4.2  / page 98 

Problems: 4E1-4E5; 4M1, 4M2, 4M3   

### 4E1. In the model definition below, which line is the likelihood?  
> $y \sim Normal(\mu, \sigma)\Leftarrow$  
> $\mu \sim Normal(0; 10)$  
> $\sigma \sim Uniform(0; 10)$  

### 4E2. In the model definition just above, how many parameters are in the posterior distribution?
> 2  

### 4E3. Using the model definition above, write down the appropriate form of Bayes’ theorem that includes the proper likelihood and priors.  
> $$Pr(\mu, \sigma | y) = \frac{\Pi_i Normal(y_i|\mu, \sigma) Normal(\mu|0; 10) Uniform(\sigma|0; 10)}{\iint \Pi_i Normal(y_i|\mu, \sigma)Normal(\mu|0; 10) Uniform(\sigma|0; 10)~d\mu~d\sigma}$$

### 4E4. In the model definition below, which line is the linear model?  
> $y_i \sim Normal(\mu, \sigma)$  
> $\mu_i = \alpha + \beta x_i \Leftarrow$  
> $\alpha \sim Normal(0; 10)$  
> $\beta \sim Normal(0; 1)$  
> $\sigma \sim Uniform(0; 10)$  

### 4E5. In the model definition just above, how many parameters are in the posterior distribution?  
> 3

### 4M1. For the model definition below, simulate observed heights from the prior (not the posterior).  
> $y_i \sim Normal(\mu; \sigma)$   
> $\mu \sim Normal(0; 10)$   
> $\sigma \sim Uniform(0; 10)$ 

```{r}
sample.mu <- rnorm( 1e4 , 0 , 10 )
sample.sigma <- runif( 1e4 , 0 , 10 )
prior.h <- rnorm( 1e4 , sample.mu , sample.sigma )

# dens(prior.h)

ggplot() + aes( prior.h ) +
      geom_density( fill="gray") +
      theme_classic()
```


### 4M2. Translate the model just above into a quap formula.  
```{r}
flist <- alist(
    height ~ dnorm( mu , sigma ) ,
    mu ~ dnorm( 0 , 10 ) ,
    sigma ~ dunif( 0 , 10 )
)

```



### 4M3. Translate the quap model formula below into a mathematical model definition.  

```{r}
# flist <- alist(
# y ~ dnorm( mu , sigma ),
# mu <- a + b*x,
# a ~ dnorm( 0 , 50 ),
# b ~ dunif( 0 , 10 ),
# sigma ~ dunif( 0 , 50 )
# )
```
$y_i \sim Normal(\mu_i, \sigma)$  
$\mu_i = \alpha + \beta(x_i - \bar{x})$  
$\alpha \sim Normal(0; 50)$  
$\beta \sim Normal(0; 10)$  
$\sigma \sim Uniform(0; 50)$  