---
title: "Chapter 5"
author: "Esteban"
date: "May 8, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(rethinking)
library(GGally)
```

## 5E1. Which of the linear models below are multiple linear regressions?  
<!-- (1) $\mu_i = \alpha + \beta x_i$   -->
(2) $\mu_i = \beta_x x_i + \beta_z z_i$  
<!-- (3) $\mu_i = \alpha + \beta(x_i - z_i)$   -->
(4) $\mu_i = \alpha + \beta_x x_i + \beta_z z_i$  

## 5E3. Write down a multiple regression to evaluate the claim: Neither amount of funding nor size
of laboratory is by itself a good predictor of time to PhD degree; but together these variables are both
positively associated with time to degree. Write down the model definition and indicate which side of
zero each slope parameter should be on.  

$$time_i \sim Normal(\mu_i, \sigma) \\
\mu_i = \alpha + \beta_f Funding_i + \beta_s Size_i \\
\alpha \sim Normal(6, 1) \\
\beta_f \sim Normal(0, 1)\\
\beta_s \sim Normal(0, 1)\\
\sigma \sim Exponential(1)\\$$

(Slope should be positive)

## 5M2. Invent your own example of a masked relationship. An outcome variable should be correlated
with both predictor variables, but in opposite directions. And the two predictor variables should be
correlated with one another. 




## 5M3. It is sometimes observed that the best predictor of fire risk is the presence of firefighters—
States and localities with many firefighters also have more fires. Presumably firefighters do not cause
fires. Nevertheless, this is not a spurious correlation. Instead fires cause firefighters. Consider the
same reversal of causal inference in the context of the divorce and marriage data. How might a high
divorce rate cause a higher marriage rate? Can you think of a way to evaluate this relationship, using
multiple regression?  
> A divorce leaves two single people to be in the pool to marry again and thus increase the marriage rate. 




## 5M4. In the divorce data, States with high numbers of Mormons (members of The Church of Jesus
Christ of Latter-day Saints, LDS) have much lower divorce rates than the regression models expected.
Find a list of LDS population by State and use those numbers as a predictor variable, predicting di-
vorce rate using marriage rate, median age at marriage, and percent LDS population (possibly stan-
dardized). You may want to consider transformations of the raw percent LDS variable.  
```{r}
data(WaffleDivorce)
d <- WaffleDivorce

d$LDSpct = c(0.78, 4.52, 6.03, 1.05, 1.92, 2.65, 0.44, 0.58, 0.40, 0.74, 0.82, 5.26, 26.02, 0.45, 0.68, 0.90, 1.31, 0.80, 0.64, 0.82, 0.73, 0.40, 0.45, 0.59, 0.72, 1.17, 4.74, 1.30, 0.65, 0.38, 3.31, 0.42, 0.85, 1.50, 0.53, 1.22, 3.66, 0.41, 0.39, 0.80, 1.21, 0.77, 1.25, 66.74, 0.74, 1.13, 3.84, 0.94, 0.46, 11.67)

# d$Marriage.std <- (d$Marriage - mean(d$Marriage)) / sd(d$Marriage)
# d$MedianAgeMarriage.std <- (d$MedianAgeMarriage - mean(d$MedianAgeMarriage)) / sd(d$MedianAgeMarriage)
# d$LDSpct.std <- (d$LDSpct - mean(d$LDSpct)) / sd(d$LDSpct)


d$L <- scale(d$LDSpct)
d$A <- scale(d$MedianAgeMarriage)
d$D <- scale(d$Divorce)
d$M <- scale(d$Marriage)

mod5M4 <- quap(
  alist(
    D ~ dnorm( mean=mu, sd=sigma ),
    mu <- a + bM*M + bA*A + bL*L,
 a ~ dnorm( 0 , 0.2 ) ,
        bM ~ dnorm( 0 , 0.5 ) ,
        bA ~ dnorm( 0 , 0.5 ) ,
        bL ~ dnorm( 0 , 0.5 ) ,
        sigma ~ dexp( 1 )
    ) , data = d )
precis(mod5M4)

plot(coeftab(mod5M4))
```



# Hard:  All three exercises below use the same data, data(foxes) (part of rethinking). 79 The urban
fox (Vulpes vulpes) is a successful exploiter of human habitat. Since urban foxes move in packs and
defend territories, data on habitat quality and population density is also included. The data frame has
five columns:
(1) group: Number of the social group the individual fox belongs to
(2) avgfood: The average amount of food available in the territory
(3) groupsize: The number of foxes in the social group
(4) area: Size of the territory
(5) weight: Body weight of the individual fox

## 5H1. Fit two bivariate Gaussian regressions, using quap: (1) body weight as a linear function of
territory size (area), and (2) body weight as a linear function of groupsize. Plot the results of these
regressions, displaying the MAP regression line and the 95% interval of the mean. Is either variable
important for predicting fox body weight?  
```{r}
library(rethinking)
data(foxes)
d2 <- foxes

# 1 territory
mod5H1 <- quap(
  alist(
    weight ~ dnorm(mean = mu, sd = sigma),
    mu <- alpha + bA*area,
    alpha ~ dnorm(4.5, 1),
    bA ~ dnorm(0, 1),
    sigma ~ dexp(1)
  ),
  data = d2
)
precis(mod5H1)

area.seq <- seq(from = 0, to = 6, length.out = 100)
mu <- link(mod5H1, data = data.frame(area = area.seq))
mu.mean <- apply(mu, 2, FUN = mean)
mu.PI <- apply(mu, 2, FUN = PI, prob = .89)

plot(weight ~ area, data = d2)
lines(x = area.seq, y = mu.mean)
shade(object = mu.PI, lim = area.seq)


#2 group size
mod5H1.2 <- quap(
  alist(
    weight ~ dnorm(mean = mu, sd = sigma),
    mu <- alpha + bG * groupsize,
    alpha ~ dnorm(4.5, 1),
    bG ~ dnorm(0, 1),
    sigma ~ dexp(1)
  ),
  data = d2
)
precis(mod5H1.2)

groupsz.seq <- seq(from = min(d2$groupsize), to = max(d2$groupsize), length.out = 100)
mu <- link(mod5H1.2, data = data.frame(groupsize = groupsz.seq))
mu.mean <- apply(mu, 2, FUN = mean)
mu.PI <- apply(mu, 2, FUN = PI, prob = .89)

plot(weight ~ groupsize, data = d2)
lines(x = groupsz.seq, y = mu.mean)
shade(object = mu.PI, lim = groupsz.seq)


```


## 5H2. Now fit a multiple linear regression with weight as the outcome and both area and groupsize
as predictor variables. Plot the predictions of the model for each predictor, holding the other predictor
constant at its mean. What does this model say about the importance of each variable? Why do you
get different results than you got in the exercise just above?  

```{r}
mod5H2 <- quap(
  alist(
    weight ~ dnorm(mean = mu, sd = sigma),
    mu <- alpha + bG * groupsize + bA * area,
    alpha ~ dnorm(4.5, 1),
    bG ~ dnorm(0, 1),
    bA ~ dnorm(0, 1),
    sigma ~ dexp(1)
  ),
  data = d2
)
precis(mod5H2)

plot(coeftab(mod5H1,mod5H1.2,mod5H2), par=c("bS","bA") )

groupsz.seq <- seq(from = min(d2$groupsize), to = max(d2$groupsize), length.out = 100)
mu <- link(mod5H2, data = data.frame(groupsize = groupsz.seq))
mu.mean <- apply(mu, 2, FUN = mean)
mu.PI <- apply(mu, 2, FUN = PI, prob = .89)

plot(weight ~ groupsize, data = d2)
lines(x = groupsz.seq, y = mu.mean)
shade(object = mu.PI, lim = groupsz.seq)
```

