---
title: "Ch8HW"
author: "Esteban"
date: "July 26, 2019"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rethinking)
library(tidyverse)
```

##8E1. For each of the causal relationships below, name a hypothetical third variable that would lead to an interaction effect.  
(1) Bread dough rises because of yeast.  
> * temperature, [sugar]   

(2) Education leads to higher income.   
> * family
* local environment (i.e., opportunities e.g. city vs. rural)  
* nutrition  
* choice of major  

(3) Gasoline makes a car go.  
> * spark plugs  

##8M3. In parts of North America, ravens depend upon wolves for their food. This is because ravens
are carnivorous but cannot usually kill or open carcasses of prey. Wolves however can and do kill
and tear open animals, and they tolerate ravens co-feeding at their kills. This species relationship
is generally described as a “species interaction.” Can you invent a hypothetical set of data on raven
population size in which this relationship would manifest as a statistical interaction? Do you think
the biological interaction could be linear? Why or why not?  

```{r}
# Number of samples
N <- 1000

# Prey
prey <- rnorm(N, mean=0, sd=1)

# Wolf
m.wolf <- 0.5*prey
wolf <- rnorm( N, mean=m.wolf, sd=1 )

# Raven
m.raven <- 0.1*prey + 0.3*wolf*prey 
raven <- rnorm( N, mean=m.raven, sd=1 )

# Make data frame
df.raven <- data.frame(raven, prey, wolf)
```

```{r}
# Raven-wolf-prey model
m8M3 <- quap(
  alist(
    raven ~ dnorm( mu, sigma ),
    mu <- a + bP*prey + bW*wolf + bPW*prey*wolf,
    a ~ dnorm( 0, 1 ),
    c( bP, bW, bPW ) ~ dnorm( 0, 1 ),
    sigma ~ dunif( 0, 3 )
  ),
  data = df.raven)

precis(m8M3)
```




##8H3. Consider again the data(rugged) data on economic development and terrain ruggedness,
examined in this chapter. One of the African countries in that example, Seychelles, is far outside
the cloud of other nations, being a rare country with both relatively high GDP and high ruggedness.
Seychelles is also unusual, in that it is a group of islands far from the coast of mainland Africa, and
its main economic activity is tourism.  

One might suspect that this one nation is exerting a strong influence on the conclusions. In
this problem, I want you to drop Seychelles from the data and re-evaluate the hypothesis that the
relationship of African economies with ruggedness is different from that on other continents.  

(a) Begin by using map to fit just the interaction model:  
$y_i \sim Normal(\mu_i, \sigma)$    
$\mu_i = \alpha + \beta_A A_i + \beta_R R_i + \beta_{AR} A_i R_i$

where y is log GDP per capita in the year 2000 (log of rgdppc_2000); A is cont_africa, the dummy
variable for being an African nation; and R is the variable rugged. Choose your own priors. Compare
the inference from this model fit to the data without Seychelles to the same model fit to the full data. Does it still seem like the effect of ruggedness depends upon continent? How much has the expected relationship changed? 

```{r}
## R code 8.1
data(rugged)
d <- rugged

# make log version of outcome
d$log_gdp <- log( d$rgdppc_2000 )

# extract countries with GDP data
dd <- d[ complete.cases(d$rgdppc_2000) , ]

# rescale variables
dd$log_gdp_std <- dd$log_gdp / mean(dd$log_gdp)
dd$rugged_std <- dd$rugged / max(dd$rugged)

# split countries into Africa and not-Africa
d.A1 <- dd[ dd$cont_africa==1 , ] # Africa
d.A0 <- dd[ dd$cont_africa==0 , ] # not Africa

########################
  # Sí Sey ------------#
m8H3 <- quap(
    alist(
        log_gdp_std ~ dnorm( mu , sigma ) ,
        mu <- a + bA*cont_africa +  bR*rugged_std + bAR*cont_africa*rugged_std ,
        a ~ dnorm( 1 , 0.1 ) ,
        c(bA, bR, bAR) ~ dnorm( 0 , 0.3 ) ,
        sigma ~ dexp( 1 )
    ) ,
    data=dd )
precis(m8H3, depth=2)

  # No Sey ------------#
dd.noSey <- dd %>% subset( country!="Seychelles" )
# dd.noSey <- dd[ dd$country!="Seychelles", ]

m8H3.noSey <- quap(
    alist(
        log_gdp_std ~ dnorm( mu , sigma ) ,
        mu <- a + bA*cont_africa +  bR*rugged_std + bAR*cont_africa*rugged_std ,
        a ~ dnorm( 1 , 0.1 ) ,
        c(bA, bR, bAR) ~ dnorm( 0 , 0.3 ) ,
        sigma ~ dexp( 1 )
    ) ,
    data=dd.noSey )
precis(m8H3.noSey, depth=2)

```
> There is less of an interaction effect, indicated by the slope bAR going from 0.25 to 0.18.  

(b) Now plot the predictions of the interaction model, with and without Seychelles. Does it still
seem like the effect of ruggedness depends upon continent? How much has the expected relationship
changed?  

```{r}
par(mfrow = c(2, 2))
####------ With Seychelles -------###
d.A1.noSey <- subset(d.A1, country!="Seychelles")

## R code 8.17
# plot Africa - cont_africa=1
plot( d.A1$rugged_std , d.A1$log_gdp_std , pch=16 , col=rangi2 ,
    xlab="ruggedness (standardized)" , ylab="log GDP (as proportion of mean)" ,
    xlim=c(0,1) )

rugged_seq <- seq( from=-0.1 , to=1.1 , length.out=30 )
mu <- link( m8H3 , data=data.frame( cont_africa=1 , rugged_std=rugged_seq ) )
mu_mean <- apply( mu , 2 , mean )
mu_ci <- apply( mu , 2 , PI , prob=0.97 )

lines( rugged_seq , mu_mean , lwd=2 )
shade( mu_ci , rugged_seq , col=col.alpha(rangi2,0.3) )
mtext("African nations")

# plot non-Africa - cont_africa=0
plot( d.A0$rugged_std , d.A0$log_gdp_std , pch=1 , col="black" ,
    xlab="ruggedness (standardized)" , ylab="log GDP (as proportion of mean)" ,
    xlim=c(0,1) )
mu <- link( m8H3 , data=data.frame( cont_africa=0 , rugged_std=rugged_seq ) )
mu_mean <- apply( mu , 2 , mean )
mu_ci <- apply( mu , 2 , PI , prob=0.97 )
lines( rugged_seq , mu_mean , lwd=2 )
shade( mu_ci , rugged_seq )
mtext("Non-African nations")


###--------- Without Seychelles ------###
d.A1.noSey <- subset( d.A1, country!="Seychelles" )

## R code 8.17
# plot Africa - cont_africa=1
plot( d.A1.noSey$rugged_std , d.A1.noSey$log_gdp_std , pch=16 , col=rangi2 ,
    xlab="ruggedness (standardized)" , ylab="log GDP (as proportion of mean)" ,
    xlim=c(0,1) )

rugged.seq <- seq( from=-0.1 , to=1.1 , length.out=30 )
mu <- link( m8H3.noSey , data=data.frame( cont_africa=1 , rugged_std=rugged_seq ) )
mu_mean <- apply( mu , 2 , mean )
mu_ci <- apply( mu , 2 , PI , prob=0.97 )

lines( rugged_seq , mu_mean , lwd=2 )
shade( mu_ci , rugged_seq , col=col.alpha(rangi2,0.3) )
mtext("African nations minus Seychelles")

# plot non-Africa - cont_africa=0
plot( d.A0$rugged_std , d.A0$log_gdp_std , pch=1 , col="black" ,
    xlab="ruggedness (standardized)" , ylab="log GDP (as proportion of mean)" ,
    xlim=c(0,1) )
mu <- link( m8H3.noSey , data=data.frame( cont_africa=0 , rugged_std=rugged_seq ) )
mu_mean <- apply( mu , 2 , mean )
mu_ci <- apply( mu , 2 , PI , prob=0.97 )
lines( rugged_seq , mu_mean , lwd=2 )
shade( mu_ci , rugged_seq )
mtext("Non-African nations (minus Seychelles in model)")
```

> The slope for African nations' gdp v. ruggedness has decreased.  


(c) Finally, conduct a model comparison analysis, using WAIC. Fit three models to the data without Seychelles:  
Model 1 : $y_i \sim Normal(\mu_i, \sigma)$  
$\mu_i = \alpha + \beta_R R_i$  
Model 2 : $y_i \sim Normal(\mu_i, \sigma)$  
$\mu_i = \alpha + \beta_A A_i + \beta_R R_i$  
Model 3 : $y_i \sim Normal(\mu_i, \sigma)$  
$\mu_i = \alpha + \beta_A A_i + \beta_R R_i + \beta_{AR} A_i R_i$  
Use whatever priors you think are sensible. Plot the model-averaged predictions of this model set.
Do your inferences differ from those in (b)? Why or why not?
```{r}
m8H3.C1 <- quap(
    alist(
        log_gdp_std ~ dnorm( mu , sigma ) ,
        mu <- a + bR*rugged_std,
        a ~ dnorm( 1 , 0.1 ) ,
        bR ~ dnorm( 0, 0.3 ) ,
        sigma ~ dexp( 1 )
    ) ,
    data=dd.noSey )

m8H3.C2 <- quap(
    alist(
        log_gdp_std ~ dnorm( mu , sigma ) ,
        mu <- a + bA*cont_africa +  bR*rugged_std,
        a ~ dnorm( 1 , 0.1 ) ,
        c(bA, bR) ~ dnorm( 0 , 0.3 ) ,
        sigma ~ dexp( 1 )
    ) ,
    data=dd.noSey )

m8H3.C3 <- quap(
    alist(
        log_gdp_std ~ dnorm( mu , sigma ) ,
        mu <- a + bA*cont_africa +  bR*rugged_std + bAR*cont_africa*rugged_std ,
        a ~ dnorm( 1 , 0.1 ) ,
        c(bA, bR, bAR) ~ dnorm( 0 , 0.3 ) ,
        sigma ~ dexp( 1 )
    ) ,
    data=dd.noSey )

compare(m8H3.C1, m8H3.C2, m8H3.C3)
```
> The model with A and R is equivalent to the interaction model.


##Plus, use the tomato.csv (attached) data set and evaluate whether hypocotyl length ("hyp") is affected by shade ("trt"), species ("species") and their interaction.
```{r}
tomato <- read.csv("Tomato.csv")
tomato <- subset(tomato, select = c(hyp, trt, species))

tomato$h <- tomato$hyp / max(tomato$hyp)
tomato$s <- tomato$species %>% as.numeric()
tomato$t <- tomato$trt %>% as.numeric() - 1

head(tomato)
str(tomato)

```

```{r}
mtomato <- quap(
    alist(
        h ~ dnorm( mu , sigma ) ,
        mu <- a[s] + b*t ,
        a[s] ~ dnorm( 0 , 1 ) ,
        b ~ dnorm( 0 , 1 ) ,
        sigma ~ dexp( 1 )
    ) ,
    data=tomato )


mtomato.int <- quap(
    alist(
        h ~ dnorm( mu , sigma ) ,
        mu <- a[s] + b[s]*t ,
        a[s] ~ dnorm( 0 , 1 ) ,
        b[s] ~ dnorm( 0 , 1 ) ,
        sigma ~ dexp( 1 )
    ) ,
    data=tomato )

precis(mtomato, depth=2)
precis(mtomato.int, depth=2)

compare(mtomato, mtomato.int)
```

###=====================================   
###******** SECOND PART  ***************   
###=====================================    

7M1, 7M2
7H1, 7H2, 7H4

##7M1. Recall the tulips example from the chapter. Suppose another set of treatments adjusted the temperature in the greenhouse over two levels: cold and hot. The data in the chapter were collected at the cold temperature. You find none of the plants grown under the hot temperature developed any blooms at all, regardless of the water and shade levels. Can you explain this result in terms of interactions between water, shade, and temperature?  
> Light and water produce blooms provided temperature doesn't rise above a threshold. 


##7M2. Can you invent a regression equation that would make the bloom size zero, whenever the temperature is hot?  
$\mu_i = \gamma_{h/c}[ \alpha + \beta_w w_i + \beta_s s_i + (\beta_w + \beta_s) w_i s_i]$  
$\gamma_h = 0$  
$\gamma_c = 1$
 

##7H1. Return to the data(tulips) example in the chapter. Now include the bed variable as a predictor in the interaction model. Don’t interact bed with the other predictors; just include it as a main effect. Note that bed is categorical. So to use it properly, you will need to either construct dummy variables or rather an index variable, as explained in Chapter ??.
```{r}
library(rethinking)
data(tulips)
d <- tulips
str(d)

d$blooms_std <- d$blooms / max(d$blooms)
d$water_cent <- d$water - mean(d$water)
d$shade_cent <- d$shade - mean(d$shade)

summary(d)
apply( d[ -1 ], 2, sd ) %>% round(2)

```
```{r}
## Modified R code 8.24
m.tulip <- quap(
    alist(
        blooms_std ~ dnorm( mu , sigma ) ,
        mu <- a[bed] + bw*water_cent + bs*shade_cent + bws*water_cent*shade_cent ,
        a[bed] ~ dnorm( 0.5 , 0.25 ) ,
        bw ~ dnorm( 0 , 0.25 ) ,
        bs ~ dnorm( 0 , 0.25 ) ,
        bws ~ dnorm( 0 , 0.25 ) ,
        sigma ~ dexp( 1 )
    ) ,
    data=d )

precis(m.tulip, depth=2)

## Is creating an ID necessary? -- NO. Factor levels already incorporated in original variable.
#
# d$bed_id <- as.numeric(d$bed)
# m.tulipID <- quap(
#     alist(
#         blooms_std ~ dnorm( mu , sigma ) ,
#         mu <- a[bed_id] + bw*water_cent + bs*shade_cent + bws*water_cent*shade_cent ,
#         a[bed_id] ~ dnorm( 0.5 , 0.25 ) ,
#         bw ~ dnorm( 0 , 0.25 ) ,
#         bs ~ dnorm( 0 , 0.25 ) ,
#         bws ~ dnorm( 0 , 0.25 ) ,
#         sigma ~ dexp( 1 )
#     ) ,
#     data=d )
# 
# precis(m.tulipID, depth=2)
```


##7H2. Use WAIC to compare the model from 7H1 to a model that omits bed. What do you infer from this comparison? Can you reconcile the WAIC results with the posterior distribution of the bed coefficients?
```{r}
## R code 8.24
m.tulip.OG <- quap(
    alist(
        blooms_std ~ dnorm( mu , sigma ) ,
        mu <- a + bw*water_cent + bs*shade_cent + bws*water_cent*shade_cent ,
        a ~ dnorm( 0.5 , 0.25 ) ,
        bw ~ dnorm( 0 , 0.25 ) ,
        bs ~ dnorm( 0 , 0.25 ) ,
        bws ~ dnorm( 0 , 0.25 ) ,
        sigma ~ dexp( 1 )
    ) ,
    data=d )

coeftab( m.tulip, m.tulip.OG )
compare( m.tulip, m.tulip.OG )
```
> The model that includes bed has a slightly lower WAIC score, although given that it is smaller than the dSE, the WAIC scores are effectively the same. Bed A seems to have overall lower blooms on average than the other two beds, which are similar to each other. When we condition on bed, giving each its own intercept, we include information which contributes to our predictive power. The other model takes an average of the intercepts for each bed. The other model expects tulip in bed A to yield less blooms on average, holding all other variables at their mean. It expects slightly more in beds B and C.   

##7H4. The values in data(nettle) are data on language diversity in 74 nations. The meaning of each column is given below.  
(1) country: Name of the country
(2) num.lang: Number of recognized languages spoken
(3) area: Area in square kilometers
(4) k.pop: Population, in thousands
(5) num.stations: Number of weather stations that provided data for the next two columns
(6) mean.growing.season: Average length of growing season, in months
(7) sd.growing.season: Standard deviation of length of growing season, in months

Use these data to evaluate the hypothesis that language diversity is partly a product of food security. The notion is that, in productive ecologies, people don’t need large social networks to buffer them against risk of food shortfalls. This means ethnic groups can be smaller and more self-sufficient, leading to more languages per capita. In contrast, in a poor ecology, there is more subsistence risk, and so human societies have adapted by building larger networks of mutual obligation to provide food insurance. This in turn creates social forces that help prevent languages from diversifying. Specifically, you will try to model the number of languages per capita as the outcome variable:

Use the logarithm of this new variable as your regression outcome. (A count model would be better
here, but you’ll learn those later, in Chapter 11.) This problem is open ended, allowing you to decide how you address the hypotheses and the uncertain advice the modeling provides. If you think you need to use WAIC anyplace, please do. If you think you need certain priors, argue for them. If you think you need to plot predictions in a certain way, please do. Just try to honestly evaluate the main effects of both mean.growing.season and sd.growing.season, as well as their two-way interaction, as outlined in parts (a), (b), and (c) below. If you are not sure which approach to use, try several.  

```{r}
data(nettle)
d <- nettle 

# R code 8.27
d$lang.per.cap <- d$num.lang / d$k.pop

# log transform lang per cap and area
d$log.lang.pcap <- log( d$lang.per.cap )
d$log.area <- log( d$area )

# center variables
d$log.lang.cent <- d$log.lang.pcap - mean( d$log.lang.pcap )
d$mean.growing.cent <- d$mean.growing.season - mean( d$mean.growing.season )
d$log.area.cent <- d$log.area - mean( d$log.area )
d$sd.growing.cent <- d$sd.growing.season - mean( d$sd.growing.season )


# mean, sd, and range
summary(d)
apply( d[-1], 2, sd ) %>% round(2)
# apply( d[-1], 2, range ) %>% diff()
```


(a) Evaluate the hypothesis that language diversity, as measured by log(lang.per.cap), is positively associated with the average length of the growing season, mean.growing.season. Consider log(area) in your regression(s) as a covariate (not an interaction). Interpret your results.  

```{r}
# qplot(mean.growing.season, log.lang.pcap, data=d)
qplot(mean.growing.cent, log.lang.cent, data=d)
qplot(sd.growing.cent, log.lang.cent, data=d)

```


```{r}
m.A <- quap(
    alist(
        log.lang.cent ~ dnorm( mu , sigma ) ,
        mu <- a + bG*mean.growing.cent,
        a ~ dnorm( 0 , 0.85 ) ,
        bG ~ dnorm( 0 , 0.4 ) ,
        sigma ~ dexp( 1 )
    ) ,
    data=d )

## Checck prior
prior <- extract.prior( m.A )

# set up the plot dimensions
plot( NULL , xlim=c( -8, 5.5 ) , ylim=c( -4, 6 ) ,
    xlab="mean growing season" , ylab="log language per capita, centered" )
abline( h=min(d$log.lang.cent) , lty=2 )
abline( h=max(d$log.lang.cent) , lty=2 )

# draw 50 lines from the prior
plot_seq <- seq( from=-8.5 , to=6 , length.out=100 )
mu <- link( m.A , post=prior , data=data.frame(mean.growing.cent=plot_seq) )
for ( i in 1:100 ) lines( plot_seq , mu[i,] , col=col.alpha("black",0.3) )
```

```{r}
m.A_area <- quap(
    alist(
        log.lang.cent ~ dnorm( mu , sigma ) ,
        mu <- a + bG*mean.growing.cent + bA*log.area.cent,
        a ~ dnorm( 0 , 0.85 ) ,
        c(bG, bA) ~  dnorm( 0 , 0.4 ) ,
        sigma ~ dexp( 1 )
    ) ,
    data=d )

coeftab( m.A, m.A_area ) 
compare( m.A, m.A_area )

m.A %>% precis()
m.A_area %>% precis()
```
> In both models, there is a slight positive association between growing season and language diversity. Area does not contribute much to WAIC in comparison with the simple model, but when included is slightly negatively associated with language diversity.

(b) Now evaluate the hypothesis that language diversity is negatively associated with the standard deviation of length of growing season, sd.growing.season. This hypothesis follows from uncertainty in harvest favoring social insurance through larger social networks and therefore fewer languages. Again, consider log(area) as a covariate (not an interaction). Interpret your results.  
```{r}
m.B <- quap(
    alist(
        log.lang.cent ~ dnorm( mu , sigma ) ,
        mu <- a + bG*mean.growing.cent + bS*sd.growing.cent, 
        a ~ dnorm( 0 , 0.85 ) ,
        c( bG, bS ) ~  dnorm( 0 , 0.4 ) ,
        sigma ~ dexp( 1 )
    ) ,
    data=d )

m.B_area <- quap(
    alist(
        log.lang.cent ~ dnorm( mu , sigma ) ,
        mu <- a + bG*mean.growing.cent + bS*sd.growing.cent + bA*log.area.cent,
        a ~ dnorm( 0 , 0.85 ) ,
        c( bG, bS, bA ) ~  dnorm( 0 , 0.4 ) ,
        sigma ~ dexp( 1 )
    ) ,
    data=d )

coeftab( m.A, m.A_area, m.B, m.B_area ) 
compare( m.A, m.A_area, m.B, m.B_area )
```
> There is a stronger negative assocaition between language diversity and the sd of the growing season.

(c) Finally, evaluate the hypothesis that mean.growing.season and sd.growing.season interact to synergistically reduce language diversity. The idea is that, in nations with longer average growing seasons, high variance makes storage and redistribution even more important than it would be otherwise. That way, people can cooperate to preserve and protect windfalls to be used during the droughts. These forces in turn may lead to greater social integration and fewer languages.]
```{r}
m.C <- quap(
    alist(
        log.lang.cent ~ dnorm( mu , sigma ) ,
        mu <- a + bG*mean.growing.cent + bS*sd.growing.cent + 
          bGS*mean.growing.cent*sd.growing.cent, 
        a ~ dnorm( 0 , 0.85 ) ,
        c( bG, bS, bGS ) ~  dnorm( 0 , 0.4 ) ,
        sigma ~ dexp( 1 )
    ) ,
    data=d )

m.C_area <- quap(
    alist(
        log.lang.cent ~ dnorm( mu , sigma ) ,
        mu <- a + bG*mean.growing.cent + bS*sd.growing.cent + 
          bA*log.area.cent + bGS*mean.growing.cent*sd.growing.cent,
        a ~ dnorm( 0 , 0.85 ) ,
        c( bG, bS, bA, bGS ) ~  dnorm( 0 , 0.4 ) ,
        sigma ~ dexp( 1 )
    ) ,
    data=d )

coeftab( m.A, m.A_area, m.B, m.B_area, m.C, m.C_area ) 
compare( m.A, m.A_area, m.B, m.B_area, m.C, m.C_area )

m.C %>% precis()

```
> The interaction model has the best WAIC score, dSE above the other models. Although growing season still has a positive interaction, both sd of growing season and the interaction with growing season have a negative impact on language diversity. [Can we say anything about sigma being smaller in this model as we do in ANOVA--that is does sigma decrease as our model encompasses more predictor variables?]
